---
title: "STAT570 HW2 Ecenaz Tunç"
output:
  html_document:
    df_print: paged
date: "2023-11-21"
---

BOOLEAN OPERATORS

First, two libraries are required for the data sets and functions.


```{r}
library(tidyverse)
library(babynames)
```

There are three main boolean operators:

1) AND &
2) OR |
3) NOT !

We want words that does not contain any vowels.
So, this is the code for that.


```{r}
str_view(words, "^[^aeiou]+$")
```



Let's break down the components of this regular expression:

The ^ at the beginning, means that it matches any character that is NOT in the specified set. 

^: Anchors in the begging ensures that the pattern starts matching from the start of the string.

[^aeiou]:  We are defining a class using []. 
In this case, the set is aeiou, which are the vowels. So, [^aeiou] matches any consonant.

 +: By using this, we are sure that the patterns matches one or more consecutive consonants.

$:This ensures that the pattern continues matching until the end of the string.

Putting it all together, ^[^aeiou]+$ makes the output string includes only consonants.
But you can make this problem a bit easier by flipping the problem around. 
Instead of looking for words that contain only consonants, we could look for words that don’t contain any vowels:


But we can make it easier by excluding the vowels:


```{r}
str_view(words[!str_detect(words, "[aeiou]")])
```



This is a useful technique whenever you’re dealing with logical combinations, particularly those involving “and” or “not”. 

For example, imagine if you want to find all words that contain “a” and “b”. 
There’s no “and” operator built in to regular expressions so we have to tackle it by looking for all words that contain an “a” followed by a “b”, or a “b” followed by an “a”:

EXAMPLE A AND . ALL WORDS THAT CONTAIN A AND B.


```{r}
str_view(words, "a.*b|b.*a")
```



#The regular expression a.*b|b.*a is used to find all words that contain both "a" and "b," regardless of the order in which they appear.
Let's break down the components of this regular expression:

a: Matches the character "a" literally.
 .*: The . matches any character, and * quantifies it to allow for zero or more occurrences.

b: Matches the character "b" literally.

|: Represents the OR operator.

b.*a: This matches any sequence that starts with "b" and is followed by zero or more characters and then an "a."

Putting it all together, the regular expression a.*b|b.*a can be read as "match a sequence that has 'a' followed by 'b' OR 'b' followed by 'a'."

The str_view function is used to visualize the matches in the specified text.
From the output you can see that <ab> is highlighted.



It’s simpler to combine the results of two calls to str_detect():


```{r}
words[str_detect(words, "a") & str_detect(words, "b")]
```



What if we wanted to see if there was a word that contains all vowels? If we did it with patterns we’d need to generate 5! (120) different patterns:

```{r}
words[str_detect(words, "a.*e.*i.*o.*u")]
# ...
words[str_detect(words, "u.*o.*i.*e.*a")]
```




It’s much simpler to combine five calls to str_detect():
  

we are attempting to find words that contain all vowels ('a', 'e', 'i', 'o', 'u') in a specific order. 
However, this approach involves generating and checking 120 different patterns, which is highly impractical and not efficient.


Here, each line represents a pattern attempting to match a word with all vowels in a specific order. 
The .* between each vowel allows for any number of characters to appear between the vowels.





So, easier way to this is with the boolean 

```{r}
words[
  str_detect(words, "a") &
    str_detect(words, "e") &
    str_detect(words, "i") &
    str_detect(words, "o") &
    str_detect(words, "u")
]
```



EXAMPLES FOR BOOLEAN OPERATORS

Simple example

In this example we can see how and or and not operators are working.

AND operator




```{r}
x <- c(TRUE, FALSE, TRUE)
y <- c(FALSE, TRUE, TRUE)
```



 Element-wise AND



```{r}
result_and <- x & y
result_and

```


 OR operator



```{r}
result_or <- x | y
result_or
```


 NOT operator


```{r}
result_not <- !x
result_not
```


SECOND EXAMPLE

Create a sample data frame

```{r}

classes <- data.frame(
  class_Name = c("stat203", "stat204", "stat361", "stat291"),
  credit = c(3, 5, 4, 4),
  capacity = c(62, 42, 50, 54)

)


# Filter higher credits and capacity
credit_capacity <- classes[classes$credit >3  & classes$capacity> 50 , ]

credit_capacity
```

In this example we can filter the data and find the classes that are match with the conditions we give.
